Collective excitations in competing phases in two and three dimensions

### 2023-12-11 paper/main.tex
ssd
asd

\newcommand{\vk}{\vec{k}}
\newcommand{\vQ}{\vec{Q}}
\newcommand{\up}{\uparrow}
\newcommand{\down}{\downarrow}
\newcommand{\kplusQ}{\vk+\vQ}
\newcommand{\kminusQ}{-\vk-\vQ}

@@

We observe that the entire Hamiltonian and thus all expectation values only depend on
\begin{equation}
    \hat{\gamma}(\vk) \coloneqq \frac{1}{D} \sum_{\alpha=1}^D \cos(k_\alpha)\,,
\end{equation}
i.e., for any operator $\hat{O}$ the following relation holds 
\begin{equation}
    \label{eqn:equal_expecs}
    \langle \hat{O}_{\vk} \rangle = \langle \hat{O}_{\vk'} \rangle \eqqcolon \langle \hat{O}( \gamma ) \rangle\,,
\end{equation}
if $\hat{\gamma}(\vk) = \hat{\gamma}(\vk')$.
While two-dimensional systems around the size of $100\times100$ lattice sites can still be solved by using momentum sums,
computing solutions for large three-dimensional systems becomes impossible due to the $N=L^3$ scaling.
This issue can be resolved by using the aforementioned fact and replacing the momentum sums with energy integrals using the bare density of states (DOS)
\begin{equation}
    \rho(\gamma) \coloneqq  \frac{1}{N} \sum_{\vk} \delta \left(\gamma - \hat{\gamma} (\vk) \right)\,.
\end{equation}
As an example, we consider 
\begin{equation}
    \Delta_\text{SC} = \frac{U}{2N} \int_{-1}^{1} \dgamma \rho(\gamma) \langle f( \gamma ) \rangle\,.
\end{equation}
We will be using the exact densities of states for the square and the simple cubic lattice found in Ref. \cite{Hanisch97}.
This allows us to access both two- and three-dimensional models on equal footing.
We can compute any phase as long as it does not introduce another kind of momentum dependence.

@@

We start with some operator set $\mathcal{B}$ that is complete with respect to commutation with a Hamiltonian $H$, 
i.e., any commutator $[H, A]$ for any $A \in \mathcal{B}$ can be represented by linear combinations of operators in $\mathcal{B}$.
We can then express any time-dependent operator from this set as
\begin{equation}
    a(t) = \sum_j c_j(t) A_j\,,
\end{equation}
where $c_j(t)$ captures the entire time dependence. 
Inserting this into the Heisenberg equation of motion and applying an operator scalar product $(A_i|\cdot)$ to both sides of the equation yields
\begin{equation}
    \label{eqn:heisenberg}
    \begin{aligned}
        \ddt a(t) = \sum_j \ddt c_j(t) A_j &= i \sum_j c_j(t) [H, A_j] \\
        \Rightarrow \sum_j \underbrace{(A_i | A_j)}_{\coloneqq \mN_{ij}} \ddt c_j(t) &= i \sum_j \underbrace{(A_i | [H, A_j])}_{\coloneqq \mM_{ij}} c_j(t) \\
        \Rightarrow \mN \ddt \vec{c}(t) &= i \mM \vec{c}(t)\,.
    \end{aligned}
\end{equation}
The matrices $\mN$ and $\mM$ contain all the dynamics and energetic properties of the system.
The advantage is that one can now handle simple matrices rather than operators acting on an enormous Hilbert space.
The size of these matrices depends entirely on the set of operators.

@@

We create our operator set on the basis of \eqref{eqn:equal_expecs}.
It consists of operators of the type
\begin{equation}
    A_\gamma \coloneqq \frac{1}{\sqrt{N}} \sum_{\vk} \delta (\gamma - \hat{\gamma}( \vk )) A_{\vk}\,,
\end{equation}
where $A_{\vk}$ represents each type of operator in \eqref{eqn:operators} and their Hermitian conjugates.
Operators with different indices are put into the set individually, e.g., $n_{\gamma \up}$, and $n_{\gamma \down}$ are distinct operators in our set.

@@

To do that, we embrace this matrix structure in \eqref{eqn:xp_set} and assume that all matrix entries are real.
Then, with $\mathcal{R}|_X$ denoting the block of a matrix $\mathcal{R}$ corresponding to the $X$-operators, we can calculate
\begin{align}
    r(z) &\coloneqq  \mathcal{R}(z)|_X = \left. \frac{1}{\mM + z \mN} \right\vert_X \nonumber \\
        &= \left[ \frac{1}{\mM} - \frac{1}{\mM} z \mN \frac{1}{\mM} + \frac{1}{\mM} z \mN \frac{1}{\mM} z \mN \frac{1}{\mM} - \cdots \right]_X \nonumber \\
        &= \left. \frac{1}{\mM} \sum_{j=0}^\infty \left( -z \mN \frac{1}{\mM} \right)^j \right\vert_X \,.
\end{align}
We may omit every second element of the sum, as we want to obtain the upper left block of the matrix, and $\mN$ always swaps the upper left and lower right block.
Thus, we obtain
\begin{align}
    r(z) &= \left. \frac{1}{\mM} \sum_{j=0}^\infty z^{2j} \left( \mN \frac{1}{\mM} \right)^{2j} \right\vert_X \nonumber \\
        &= \left. \frac{1}{\mM - z^2 \mN \mM^{-1} \mN} \right\vert_X \nonumber \\
        &= \frac{1}{\mathcal{K}_+ - z^2 \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger}\,.
\end{align}
Next, we define $\check{N} \coloneqq \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger$ for brevity.
Since $\mathcal{M}$ is positive semidefinite and block diagonal, its blocks $\mathcal{K}_\pm$ must be positive semidefinite as well.
Therefore, $\check{N}$ is also positive semidefinite and we can define its square root.
We obtain
\begin{equation}
    r(z) = \check{N}^{-1/2} \frac{1}{\check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2} - z^2} \check{N}^{-1/2}\,.
\end{equation}
The matrix $\check{M} \coloneqq  \check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2}$ is positive semidefinite and Hermitian by the previous arguments.
If $\check{N}$ has a singular part, the inverse can be replaced by the pseudo-inverse.
The same calculation can be repeated by exchanging $X$ with $P$, yielding the relation for the remaining block.

@@

\begin{widetext}
\begin{equation}
    \left( z^2 - \begin{pmatrix}
        a_0 & b_1 & 0 & 0 & \cdots \\
        b_1 & a_1 & b_2 & 0 & \cdots \\
        0 & b_2 & a_2 & b_3 & \cdots \\
        0 & 0 & b_3 & a_3 & \ddots \\
        \vdots & \vdots & \vdots & \ddots & \ddots
    \end{pmatrix} \right)_{00}^{-1} = \dfrac{1}{z^2 - a_0 - \dfrac{b_1^2}{z^2 - a_1 - \dfrac{b_2^2}{ z^2 - a_2 - \hdots}}}\,\,
\end{equation}
\end{widetext}

where $a_i$ and $b_i$ are the Lanczos coefficients \cite{PettiforRecursion,ViswanathRecursion}.
In the single-band case, the coefficients approach the limit
\begin{equation}
    \label{eqn:inf_lanczos}
    a_\infty = \frac{\omega_+ + \omega_-}{2}\text{  and  } b_\infty = \frac{\omega_+ - \omega_-}{4}\,,
\end{equation}
where $\omega_\pm$ represents the upper and lower band edge, respectively \cite{PettiforRecursion}.
This behaviour allows us to terminate the continued fraction, once the coefficients have converged close enough to these values, using the so-called square root terminator
\begin{equation}
    T(\omega) = \frac{1}{2b_\infty^2} \left( \omega - a_\infty \mp \sqrt{(\omega - a_\infty)^2 - 4 b_\infty^2} \right)\,,
\end{equation}
where the negative sign is to be chosen for $\omega - a_\infty > 2b_\infty$ and $|\omega - a_\infty| < 2b_\infty$ and the positive one otherwise.
It is to be noted, that the coefficients start to drop off from this limit as more and more Lanczos iterations are performed \cite{ViswanathRecursion}.
Therefore, we terminate the continued fraction with the aforementioned terminator at the coefficient pair that has the least deviation from $a_\infty$ and $b_\infty$.

@@

In this section, we will be investigating four different diagonal Green's functions $G_{AA^\dagger}(\omega + i0^+)$, namely with
\begin{subequations}
    \label{eqn:resolvent_bases}
    \begin{align}
        A_\text{Higgs} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} + f_{\vk}^\dagger \right) \nonumber \\ 
            &= \int_{-1}^{1} \rho(\gamma) \left( f(\gamma) + f^\dagger (\gamma) \right) \mathrm{d}\gamma \,,\\
        A_\text{Phase} &= \frac{i}{\sqrt{N}} \sum_{\vk} \left( f_{\vk} - f_{\vk}^\dagger \right) \nonumber \\ 
            &= i \int_{-1}^{1} \rho(\gamma) \left( f(\gamma) - f^\dagger (\gamma) \right) \mathrm{d}\gamma \,,\\
        A_\text{CDW} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} + g_{\vk \down} \right) \nonumber \\ 
            &= \int_{-1}^{1} \rho(\gamma) \left( g_{\up}(\gamma) + g_{\down}(\gamma) \right) \mathrm{d}\gamma \,,\\
        A_\text{AFM} &= \frac{1}{\sqrt{N}} \sum_{\vk} \left( g_{\vk \up} - g_{\vk \down} \right) \nonumber \\ 
            &= \int_{-1}^{1} \rho(\gamma) \left( g_{\up}(\gamma) - g_{\down}(\gamma) \right) \mathrm{d}\gamma \,,
    \end{align}
\end{subequations}
where $N$ is the number of lattice sites in the system. 
Each one of these operators generates a different kind of collective mode.
The first operator captures the amplitude mode of the $s$-wave superconducting state, while the second one catches the phase mode \cite{Fan22}.
The remaining two capture the collective behaviour of the CDW and AFM states, respectively.
Note, that this choice of normalization cancels out with other system size-dependent terms so that our final results do not depend on it at all.

@@

\appendix
\section{Obtaining Green's functions from the iEoM-matrices}

Nevertheless, let us show that $\mathcal{M}$ is positive semidefinite if the system is in thermal equilibrium.
For any $\vec{x} \in \mathbb{C}^N$, consider
\begin{align}
    \vec{x}^\dagger \mM \vec{x} &= \sum_{ij} x_i^* x_j \mM_{ij} \nonumber \\
        &= \langle \left[ \sum_i x_i^* A_i^\dagger, \left[ H, \sum_j x_j A_j \right] \right]  \rangle \nonumber \\
        &= \langle [B^\dagger, [H, B]] \rangle\,,\quad B \coloneqq  \sum_j x_j A_j\,.
\end{align}
Expanding the thermal expectation value $\langle O \rangle = \frac{1}{Z} \Trace [\exp(-\beta H_\text{MF}) O]$, where $\beta$ is the inverse temperature and $Z$ the partition function, and the commutators, we obtain

\begin{widetext}
    \begin{align}
        \langle [B^\dagger, [H, B]] \rangle &= \frac{1}{Z} \Trace [\exp(-\beta H) [B^\dagger, [H, B]]] \nonumber \\
            &= \frac{1}{Z} \sum_j e^{-\beta \epsilon_j} \bra{\psi_j} [B^\dagger, [H, B]] \ket{\psi_j} \nonumber \\
            &= \frac{1}{Z} \sum_j e^{-\beta \epsilon_j} \left[ \bra{\psi_j} B^\dagger (H - \epsilon_j) B \ket{\psi_j} + \bra{\psi_j} B (H - \epsilon_j) B^\dagger \ket{\psi_j} \right] \nonumber \\
            &= \frac{1}{Z} \sum_{ij} e^{-\beta \epsilon_j} \left[ \bra{\psi_j} B^\dagger \ket{\psi_i} (\epsilon_i - \epsilon_j) \bra{\psi_i} B \ket{\psi_j} + \bra{\psi_j} B \ket{\psi_i} (\epsilon_i - \epsilon_j) \bra{\psi_i} B^\dagger \ket{\psi_j} \right] \nonumber \\
            &= \frac{1}{Z} \sum_{ij} e^{-\beta \epsilon_j} (\epsilon_i - \epsilon_j) \underbrace{ \left[ \bra{\psi_j} B^\dagger \ket{\psi_i} \bra{\psi_i} B \ket{\psi_j} + \bra{\psi_j} B \ket{\psi_i} \bra{\psi_i} B^\dagger \ket{\psi_j} \right]}_{\coloneqq C_{ij} = C_{ji} \geq 0} \nonumber \\
            &= \frac{1}{2Z} \left[ \sum_{ij} C_{ij} e^{-\beta \epsilon_j} (\epsilon_i - \epsilon_j) - \sum_{ij} C_{ij} e^{-\beta \epsilon_i} (\epsilon_i - \epsilon_j) \right] \nonumber \\
            &= \frac{1}{2Z} \sum_{ij} C_{ij} \left( e^{-\beta \epsilon_j} - e^{-\beta \epsilon_i} \right) \left( \epsilon_i - \epsilon_j \right) \geq 0\,.
    \end{align}
\end{widetext}

Thus, we are able to compute a real square root of $\mathcal{M}$.

>>>>

\newcommand{\vk}{\vec{k}}
\newcommand{\vl}{\vec{l}}
\newcommand{\vQ}{\vec{Q}}
\newcommand{\up}{\uparrow}
\newcommand{\down}{\downarrow}
\newcommand{\kplusQ}{\vk+\vQ}
\newcommand{\kminusQ}{-\vk-\vQ}

@@

We observe that the entire Hamiltonian and thus all expectation values only depend on
\begin{equation}
    \hat{\gamma}(\vk) \coloneqq \frac{1}{D} \sum_{\alpha=1}^D \cos(k_\alpha)\,,
\end{equation}
i.e., for any operator $\hat{O}$ the following relation holds 
\begin{equation}
    \label{eqn:equal_expecs}
    \langle \hat{O}_{\vk} \rangle = \langle \hat{O}_{\vk'} \rangle \eqqcolon \langle \hat{O}( \gamma ) \rangle\,,
\end{equation}
if $\hat{\gamma}(\vk) = \hat{\gamma}(\vk')$.
While two-dimensional systems around the size of $100\times100$ lattice sites can still be solved by using momentum sums,
computing solutions for large three-dimensional systems becomes impossible due to the $N=L^3$ scaling.
This issue can be resolved by using the aforementioned fact and replacing the momentum sums with energy integrals using the bare density of states (DOS)
\begin{equation}
    \rho(\gamma) \coloneqq  \frac{1}{N} \sum_{\vk} \delta \left(\gamma - \hat{\gamma} (\vk) \right)\,.
\end{equation}
As an example, we consider 
\begin{equation}
    \Delta_\text{SC} = \frac{U}{N} \int_{-1}^{1} \dgamma \rho(\gamma) \langle f( \gamma ) \rangle\,.
\end{equation}
We will be using the exact densities of states for the square and the simple cubic lattice found in Ref. \cite{Hanisch97}.
This allows us to access both two- and three-dimensional models on equal footing.
We can compute any phase as long as it does not introduce another kind of momentum dependence.

@@

We start with some operator set $\mathcal{B}$ that is complete with respect to commutation with a Hamiltonian $H$, 
i.e., any commutator $[H, A]$ for any $A \in \mathcal{B}$ can be represented by linear combinations of operators in $\mathcal{B}$.
We can then express any time-dependent operator from this set as
\begin{equation}
    a(t) = \sum_j c_j(t) A_j\,,
\end{equation}
where $c_j(t)$ captures the entire time dependence. 
Inserting this into the Heisenberg equation of motion and applying an operator scalar product $(A_i|\cdot)$ to both sides of the equation yields
\begin{equation}
    \label{eqn:heisenberg}
    \begin{aligned}
        \ddt a(t) = \sum_j \ddt c_j(t) A_j &= i \sum_j c_j(t) [H, A_j] \\
        \Rightarrow \sum_j \underbrace{(A_i | A_j)}_{\coloneqq \mN_{ij}} \ddt c_j(t) &= i \sum_j \underbrace{(A_i | [H, A_j])}_{\coloneqq \mM_{ij}} c_j(t) \\
        \Rightarrow \mN \ddt \vec{c}(t) &= i \mM \vec{c}(t)\,.
    \end{aligned}
\end{equation}
The matrices $\mN$ and $\mM$ contain all the dynamics and energetic properties of the system.
The former is referred to as a dynamical matrix while the latter is called a norm matrix.
The advantage is that one can now handle simple matrices rather than operators acting on an enormous Hilbert space.
The size of these matrices depends entirely on the set of operators.

@@

We create our operator set on the basis of \eqref{eqn:equal_expecs}.
It consists of operators of the type
\begin{equation}
    \label{eqn:ieom_basis_operator}
    A_\gamma \coloneqq \frac{1}{\sqrt{N}} \sum_{\vk} \delta (\gamma - \hat{\gamma}( \vk )) A_{\vk}\,,
\end{equation}
where $A_{\vk}$ represents each type of operator in \eqref{eqn:operators} and their Hermitian conjugates.
Operators with different indices are put into the set individually, e.g., $n_{\gamma \up}$, and $n_{\gamma \down}$ are distinct operators in our set.

@@

To do that, we embrace this matrix structure in \eqref{eqn:xp_set} and assume that all matrix entries are real.
Then, with $\mathcal{R}|_X$ denoting the block of a matrix $\mathcal{R}$ corresponding to the $X$-operators, we can calculate
\begin{align}
    r_X (z) &\coloneqq \mathcal{R}(z)|_X = \left. \frac{1}{\mM + z \mN} \right\vert_X \nonumber \\
        &= \left[ \frac{1}{\mM} - \frac{1}{\mM} z \mN \frac{1}{\mM} + \frac{1}{\mM} z \mN \frac{1}{\mM} z \mN \frac{1}{\mM} - \cdots \right]_X \nonumber \\
        &= \left. \frac{1}{\mM} \sum_{j=0}^\infty \left( -z \mN \frac{1}{\mM} \right)^j \right\vert_X \,.
\end{align}
We may omit every second element of the sum, as we want to obtain the upper left block of the matrix, and $\mN$ always swaps the upper left and lower right block.
Thus, we obtain
\begin{align}
    r_X (z) &= \left. \frac{1}{\mM} \sum_{j=0}^\infty z^{2j} \left( \mN \frac{1}{\mM} \right)^{2j} \right\vert_X \nonumber \\
        &= \left. \frac{1}{\mM - z^2 \mN \mM^{-1} \mN} \right\vert_X \nonumber \\
        &= \frac{1}{\mathcal{K}_+ - z^2 \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger}\,.
\end{align}
Next, we define $\check{N} \coloneqq \mathcal{L} \mathcal{K}_-^{-1} \mathcal{L}^\dagger$ for brevity.
Since $\mathcal{M}$ is positive semidefinite and block diagonal, its blocks $\mathcal{K}_\pm$ must be positive semidefinite as well.
Therefore, $\check{N}$ is also positive semidefinite and we can define its square root.
We obtain
\begin{equation}
    r_X (z) = -\check{N}^{-1/2} \frac{1}{z^2 - \check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2}} \check{N}^{-1/2}\,.
\end{equation}
The matrix $\check{M} \coloneqq \check{N}^{-1/2} \mathcal{K}_+ \check{N}^{-1/2}$ is positive semidefinite and Hermitian by the previous arguments.
If $\check{N}$ has a singular part, the inverse can be replaced by the pseudo-inverse.
The same calculation can be repeated by exchanging $X$ with $P$, yielding the relation for the remaining block.
Then, the definitions of the matrices above change in the following way
\begin{equation}
    \check{N}_P \coloneqq \mathcal{L}^\dagger \mathcal{K}_+^{-1} \mathcal{L}\,,\quad \check{M}_P \coloneqq \check{N}_P^{-1/2} \mathcal{K}_- \check{N}_P^{-1/2}\,.
\end{equation}
In conclusion, the Green's function is given by
\begin{subequations}
    \begin{align}
        G_{AB^\dagger} \vert_P (z) &= - \left[ \vec{b}^\dagger (0) \right]_P \mathcal{L}^\dagger r_X (z) \mathcal{L} \left[ \vec{a}(0) \right]_P\,, \\
        G_{AB^\dagger} \vert_X (z) &= - \left[ \vec{b}^\dagger (0) \right]_X \mathcal{L} r_P (z) \mathcal{L}^\dagger \left[ \vec{a}(0) \right]_X\,.
    \end{align}
\end{subequations}

@@

\begin{widetext}
\begin{equation}
    \left( z^2 - \begin{pmatrix}
        a_0 & b_1 & 0 & 0 & \cdots \\
        b_1 & a_1 & b_2 & 0 & \cdots \\
        0 & b_2 & a_2 & b_3 & \cdots \\
        0 & 0 & b_3 & a_3 & \cdots \\
        \vdots & \vdots & \vdots & \vdots & \ddots
    \end{pmatrix} \right)_{00}^{-1} = \dfrac{1}{z^2 - a_0 - \dfrac{b_1^2}{z^2 - a_1 - \dfrac{b_2^2}{ z^2 - a_2 - \hdots}}}\,\,
\end{equation}
\end{widetext}

where $a_i$ and $b_i$ are the Lanczos coefficients \cite{PettiforRecursion,ViswanathRecursion}.
Obtaining a diagonal block of $\mathcal{G}$ is easily achieved by setting $\vec{b}(0) = \vec{a}(0)$.
Obtaining an off-diagonal block requires linear combinations in the form of \textcolor{red}{Ich bin damit nicht glücklich, da $\mathcal{G}$ ja eigentlich nicht hermitisch ist (wegen dem $z = \omega + i0^+$).
Das ist für die Ergebnisse völlig egal, da ich eh nur diagonale Blöcke betrachte, aber nunja...}
\begin{subequations}
    \begin{align}
        4 \Re [ \vec{a}^\dagger \mathcal{G} \vec{b} ] &= (\vec{a} + \vec{b})^\dagger \mathcal{G} (\vec{a} + \vec{b}) - (\vec{a} - \vec{b})^\dagger \mathcal{G} (\vec{a} - \vec{b}) \nonumber \\
            &= 2 \vec{a}^\dagger \mathcal{G} \vec{b} + 2 \vec{b}^\dagger \mathcal{G} \vec{a}\,, \\
        4i \Im [ \vec{a}^\dagger \mathcal{G} \vec{b} ] &= (\vec{a} + i \vec{b})^\dagger \mathcal{G} (\vec{a} + i \vec{b}) - (\vec{a} - i \vec{b})^\dagger \mathcal{G} (\vec{a} - i \vec{b}) \nonumber \\
            &= 2i \vec{a}^\dagger \mathcal{G} \vec{b} - 2i \vec{b}^\dagger \mathcal{G} \vec{a}\,.
    \end{align}
\end{subequations}
In the single-band case, the coefficients approach the limit
\begin{equation}
    \label{eqn:inf_lanczos}
    a_\infty = \frac{\omega_+ + \omega_-}{2}\text{  and  } b_\infty = \frac{\omega_+ - \omega_-}{4}\,,
\end{equation}
where $\omega_\pm$ represents the upper and lower band edge, respectively \cite{PettiforRecursion}.
This behaviour allows us to terminate the continued fraction, once the coefficients have converged close enough to these values, using the so-called square root terminator
\begin{equation}
    T(\omega) = \frac{1}{2b_\infty^2} \left( \omega - a_\infty \mp \sqrt{(\omega - a_\infty)^2 - 4 b_\infty^2} \right)\,,
\end{equation}
where the negative sign is to be chosen for $\omega - a_\infty > 2b_\infty$ and $|\omega - a_\infty| < 2b_\infty$ and the positive one otherwise.
It is to be noted, that the coefficients start to drop off from this limit as more and more Lanczos iterations are performed \cite{ViswanathRecursion}.
Therefore, we terminate the continued fraction with the aforementioned terminator at the coefficient pair that has the least deviation from $a_\infty$ and $b_\infty$.

@@

In this section, we will be investigating four different diagonal Green's functions $G_{AA^\dagger}(\omega + i0^+)$, namely with
\begin{subequations}
    \label{eqn:resolvent_bases}
    \begin{align}
        A_\text{Higgs} &= \frac{1}{N} \sum_{\vk} \left( f_{\vk} + f_{\vk}^\dagger \right) \,,\\
        A_\text{Phase} &= \frac{i}{N} \sum_{\vk} \left( f_{\vk} - f_{\vk}^\dagger \right) \,,\\
        A_\text{CDW} &= \frac{1}{N} \sum_{\vk} \left( g_{\vk \up} + g_{\vk \down} \right) \,,\\
        A_\text{AFM} &= \frac{1}{N} \sum_{\vk} \left( g_{\vk \up} - g_{\vk \down} \right) \,,
    \end{align}
\end{subequations}
where $N$ is the number of lattice sites in the system. 
Each one of these operators generates a different kind of collective mode.
The first operator captures the amplitude mode of the $s$-wave superconducting state, while the second one catches the phase mode \cite{Fan22}.
The remaining two capture the collective behaviour of the CDW and AFM states, respectively.
Note, that this choice of normalization cancels out with other system size-dependent terms so that our final results do not depend on it at all.

@@

\appendix
\section{Semipositivity of the dynamical matrix}

We show that $\mathcal{M}$ is positive semidefinite if the system is in thermal equilibrium.
For any $\vec{x} \in \mathbb{C}^N$, consider
\begin{align}
    \vec{x}^\dagger \mM \vec{x} &= \sum_{ij} x_i^* x_j \mM_{ij} \nonumber \\
        &= \langle \left[ \sum_i x_i^* A_i^\dagger, \left[ H, \sum_j x_j A_j \right] \right]  \rangle \nonumber \\
        &= \langle [B^\dagger, [H, B]] \rangle\,,\quad B \coloneqq  \sum_j x_j A_j\,.
\end{align}
Expanding the thermal expectation value $\langle O \rangle = \frac{1}{Z} \Trace [\exp(-\beta H_\text{MF}) O]$, where $\beta$ is the inverse temperature and $Z$ the partition function, and the commutators, we obtain

\begin{widetext}
    \begin{align}
        \langle [B^\dagger, [H, B]] \rangle &= \frac{1}{Z} \Trace [\exp(-\beta H) [B^\dagger, [H, B]]] \nonumber \\
            &= \frac{1}{Z} \sum_j e^{-\beta \epsilon_j} \bra{\psi_j} [B^\dagger, [H, B]] \ket{\psi_j} \nonumber \\
            &= \frac{1}{Z} \sum_j e^{-\beta \epsilon_j} \left[ \bra{\psi_j} B^\dagger (H - \epsilon_j) B \ket{\psi_j} + \bra{\psi_j} B (H - \epsilon_j) B^\dagger \ket{\psi_j} \right] \nonumber \\
            &= \frac{1}{Z} \sum_{ij} e^{-\beta \epsilon_j} \left[ \bra{\psi_j} B^\dagger \ket{\psi_i} (\epsilon_i - \epsilon_j) \bra{\psi_i} B \ket{\psi_j} + \bra{\psi_j} B \ket{\psi_i} (\epsilon_i - \epsilon_j) \bra{\psi_i} B^\dagger \ket{\psi_j} \right] \nonumber \\
            &= \frac{1}{Z} \sum_{ij} e^{-\beta \epsilon_j} (\epsilon_i - \epsilon_j) \underbrace{ \left[ \bra{\psi_j} B^\dagger \ket{\psi_i} \bra{\psi_i} B \ket{\psi_j} + \bra{\psi_j} B \ket{\psi_i} \bra{\psi_i} B^\dagger \ket{\psi_j} \right]}_{\coloneqq C_{ij} = C_{ji} \geq 0} \nonumber \\
            &= \frac{1}{2Z} \left[ \sum_{ij} C_{ij} e^{-\beta \epsilon_j} (\epsilon_i - \epsilon_j) - \sum_{ij} C_{ij} e^{-\beta \epsilon_i} (\epsilon_i - \epsilon_j) \right] \nonumber \\
            &= \frac{1}{2Z} \sum_{ij} C_{ij} \left( e^{-\beta \epsilon_j} - e^{-\beta \epsilon_i} \right) \left( \epsilon_i - \epsilon_j \right) \geq 0 \,.
    \end{align}
\end{widetext}

\section{Numerical treatment of the iEoM matrices}

We would like to use operators like \eqref{eqn:ieom_basis_operator} to span our operator set.
However, numerically, it is impossible to matrices with infinite dimensions, requiring us to discretize.
Additionally, we cannot numerically write down $\delta$ distributions as matrix elements. Therefore, we need some special treatments, which we will describe in this section.
We discretize the $\gamma$-space equally with each abscissa $\gamma_i$ being spaced by $\Delta \gamma$.
Then, we define an approximate $\delta$ function
\begin{equation}
    h(\gamma) \coloneqq \begin{cases}
        \frac{1}{\Delta \gamma} & |\gamma| < \frac{\Delta \gamma}{2} \\ 0 & \text{otherwise}
    \end{cases}\,.
\end{equation}
Computing an integral yields
\begin{equation}
    \int f(\tilde{\gamma}) h(\tilde{\gamma} - \gamma_j) \mathrm{d} \tilde{\gamma} \approx \sum_i f(\gamma_i) h(\gamma_i - \gamma_j) \Delta \gamma = f(\gamma_j)\,,
\end{equation}
and thereby mimicking the continuous $\delta$ distribution for a discrete description.
Note, that if one takes the limit $\Delta \gamma \to 0$ there $\delta$ distribution is exactly reproduced.

Numerically, we can compute the matrix elements like this
\begin{align}
    (A_i | A_j) &= \frac{(\Delta \gamma)^2}{N} \sum_{\vk \vl} h(\gamma_i - \hat{\gamma}(\vk)) h(\gamma_j - \hat{\gamma}(\vl)) (A_{\vk} | A_{\vl}) 
\end{align}
